# **Botulism Data**
#### The data set selected for this exercise was retrieved from the [***CDC***](https://data.cdc.gov/) data page and contains historic data on Botulism cases throughout the United States by year, case count, "BotType" (source), and "ToxinType" (strain). The raw data contains 2280 observations of 5 variables. Instead of utilizing an "NA" for missing data, "Unknown" was used. For the purpose of this exercise, I altered the "Unknown" character to reflect "NA" and omitted the NAs (a total of 403 rows were omitted). I also standardized the column name "BotType" and "ToxinType" to "Source" and "Strain", respectively. To parce the data down even further, I selected two states to evalutate; California and Georgia. 

## **Loading packages** 
```{r}
library(readr)
library(tidyverse)
```

## **Loading data into R**
```{r}
botulism <- read_csv("dataanalysis-exercise/rawdata/Botulism.csv")
```

## **Exploring Botulism data**
```{r}
str(botulism)
summary(botulism)
class(botulism)
#Overall, the data loaded into R is fairly tidy
```

## **Replacing "Unknowns" with NAs**
```{r}
botulism [botulism == "Unknown"] <- NA
str(botulism)
```
## **Removing NAs from data set**
```{r}
botulism_na<-na.omit(botulism)
```

## **Renaming columns**
```{r}
botulism_na <- botulism_na %>%
  rename("Source"="BotType",
         "Strain"="ToxinType")
```


## **Selecting data by state (CA and GA)**
```{r}
condensed_bot<- dplyr::filter(botulism_na, State %in% 
                                c("California", "Georgia"))

summary(condensed_bot)
```

## **Saving as RDS**
```{r}
saveRDS(condensed_bot, file="dataanalysis-exercise/Data/Clean Data/Botulism.RDS")
```

## **Saving summary table as RDS**
```{r}
sumtab_bot= data.frame(do.call(cbind, lapply(condensed_bot, summary)))
print(sumtab_bot)

saveRDS(sumtab_bot, file= "dataanalysis-exercise/Data/Summary Table/botsumtable.RDS")
```

