# **Botulism Data**
#### The data set selected for this exercise was retrieved from the [***CDC***](https://data.cdc.gov/) data page and contains historic data on Botulism cases throughout the United States by year, case count, "BotType" (source), and "ToxinType" (strain). The raw data contains 2280 observations of 5 variables. Instead of utilizing an "NA" for missing data, "Unknown" was used. For the purpose of this exercise, I altered the "Unknown" character to reflect "NA" and omitted the NAs (a total of 403 rows were omitted). I also standardized the column name "BotType" and "ToxinType" to "Source" and "Strain", respectively. To parce the data down even further, I selected two states to evalutate; California and Georgia. 

## **Loading packages** 
```{r}
library(readr)
library(tidyverse)
```

## **Loading data into R**
```{r}
botulism <- read_csv("dataanalysis-exercise/rawdata/Botulism.csv")
```

## **Exploring Botulism data**
```{r}
str(botulism)
summary(botulism)
class(botulism)
#Overall, the data loaded into R is fairly tidy
```

## **Replacing "Unknowns" with NAs**
```{r}
botulism [botulism == "Unknown"] <- NA
str(botulism)
```
## **Removing NAs from data set**
```{r}
botulism_na<-na.omit(botulism)
```

## **Renaming columns**
```{r}
botulism_na <- botulism_na %>%
  rename("Source"="BotType",
         "Strain"="ToxinType")
```


## **Selecting data by state (CA and GA)**
```{r}
condensed_bot<- dplyr::filter(botulism_na, State %in% 
                                c("California", "Georgia"))

summary(condensed_bot)
```

## **Saving as RDS**
```{r}
saveRDS(condensed_bot, file="dataanalysis-exercise/Data/Clean Data/Botulism.RDS")
```

## **Saving summary table as RDS**
```{r}
sumtab_bot= data.frame(do.call(cbind, lapply(condensed_bot, summary)))
print(sumtab_bot)

saveRDS(sumtab_bot, file= "dataanalysis-exercise/Data/Summary Table/botsumtable.RDS")
```

# This section added by Nathan Greenslit


## **Load Data**
```{r}
data<- readRDS("dataanalysis-exercise/Data/Clean Data/Botulism.RDS") #Loading in condensed_bot data from Kim
```

## **Wrangle Data**
```{r}
data2<- data %>%
  select(Year, Count,State, Source) #Getting rid of Strains

case_tot<- data2 %>% #This creates a column with the total counts per year instead of separating it by strain. This omits the issue of having multiples of the same year for counts. 
  group_by(Year, State, Source) %>%
  summarize_if(is.numeric, sum) %>%
  ungroup()
```

## **Create California and Georgia Specific Dataframes**
```{r}
ga<- case_tot %>%
  filter(State %in% "Georgia")

ca<- case_tot %>%
  filter(State %in% "California")


```

## **Botulism Cases By State and Source**
```{r}
case_tot %>% ggplot() +geom_line(
  aes(x = Year,
      y = Count,
      color = Source,
      linetype = State)) +
  theme_bw() +
  labs(x = "Year",
       y = "Case Counts",
       title = "Botulism Cases (1916-2017)") +
  theme(plot.title = element_text(hjust = 0.5))
```
*California appears to have a wider range of data collected (across years and different sources). Let's focus on this State*

## **Let's look at Botulism counts in California by Source**
```{r}
ca %>% ggplot() +geom_line(
  aes(x = Year,
      y = Count,
      color = Source)) +
  theme_bw() +
  labs(x = "Year",
       y = "Case Counts",
       title = "Botulism Cases in California (1916-2017)") +
  theme(plot.title = element_text(hjust = 0.5))
```
*It wasn't until the 1970's that other sources of botulism, such as infant cases, were being detected (Rosow,2015) Let's look at 1980-2020*

## **Botulism Cases in California (1980-2017)**
```{r}
ca %>% filter(Year %in% (1980:2020)) %>%
  
ggplot() +geom_boxplot(
  aes(x = Source,
      y = Count,
      color = Source)) +
  theme_bw() +
  labs(y = "Case Counts",
       title = "Botulism Cases in California (1980-2017)") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")
```
*Infant cases seem to be the most common. Looking into this, infants are at a higher risk due to their weakened immune system, lack of gastric acidity, and a diminished bacterial flora(Van Horn, 2022). Let's go back to the strain data and see which strains are most common in infants*
 
## **Infant Botulism Cases in California by Strain (1980-2017)**
```{r}
data %>% filter(Year %in% (1980:2020),
                Source %in% "Infant") %>% #Taking original dataset and filtering for 1980-2017 and for infant sources
  
ggplot() +geom_boxplot(
  aes(x = Strain,
      y = Count,
      color = Strain)) +
  theme_bw() +
  labs(y = "Case Counts",
       title = "Infant Botulism Cases in California by Strain (1980-2017)") +
  theme(plot.title = element_text(hjust = 0.5),
       legend.position = "none")
```
**Strain A seems to be the most prevalent in infants followed by Strain B. This can be confirmed at https://www.infantbotulism.org/readings/ib_chapter_6th_edition.pdf **

